---
title: "Data Manipulation"
output: html_document
---

```{r}
library(tidyverse)

```

Import data

```{r}
litters_df =
  read_csv("data/FAS_litters.csv", na = c("NA", "", " ", "."))
litters_df =
  janitor::clean_names(litters_df)


pups_df =
  read_csv("data/FAS_pups.csv", na = c("NA", "", " "), skip = 3)
pups_df =
  janitor::clean_names(pups_df)

```

## select

'select' selects variables from a data-frame

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

select a range

```{r}
select(litters_df, group:gd_of_birth)
```

select by removal

```{r}
select(litters_df, group)
select(litters_df, (group:gd_of_birth))
```

select by prefix

```{r}
select(litters_df, group, starts_with("gd"))
```

rearrange by selecting

```{r}
select(litters_df, starts_with("gd"), group)
select(litters_df, litter_number, everything())
```

rename by selecting

```{r}
select(litters_df, GROUP = group, everything())
```

What if I only select one variable?

```{r}
select(litters_df, group)
```

A Data-frame is returned!

## Learning Assessment

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## Filter

use filter() to filter out rows

```{r}
filter(litters_df, gd_of_birth == 20, pups_born_alive == 5)
```

Look for more than 5 pups alive

```{r}
filter(litters_df, pups_born_alive >= 5)
```

look for non-matches

```{r}
filter(litters_df, pups_born_alive != 5)
```

match to a set

```{r}
filter(litters_df, group %in% c("Con7", "Con8"))
```

filter out missing rows with drop_na()

```{r}
drop_na(litters_df)
drop_na(pups_df)
```

## Learning Assessment

```{r}
filter(pups_df, sex == 1)
filter(pups_df, sex == 2, pd_walk <= 11)
```

## Mutate

use mutate() to create and modify variables

```{r}
mutate(
  litters_df,
  gd_weight = gd18_weight - gd0_weight
)
```

## Pipes

Learning Assessment

```{r}
pups_df =
  read_csv("data/FAS_pups.csv", na = c("NA", "", " ", "."), skip = 3) |>
  janitor::clean_names() |>
  filter(sex == 1) |>
  select(- pd_ears) |>
  mutate(pd_pivot_ge7 = pd_pivot >= 7)

head(pups_df)
```

